Covariates analysis of MSBB RNASeq data

### 2015-12-31 code/Rmd/MSSM_RNASeq.Rmd
Major modifications in MSBB RNASeq analysis

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('Rmd/',thisFileName))
```
Data were obtained from AMP-AD portal of MSSM, and following brain regions were analysed for differential expression between case and control:
  | Region | Name                      | 
  | :----- | ------------------------- | 
  | BM_10  | Frontal Pole (FP)         |
  | BM_22  | Superior Temporal Gyrus   |
  | BM_36  | Parahipocampal Gyrus (PHG)|

@@

Remove genes that have less than 1 cpm counts in at least 50% of samples.
```{r cpmnormalisation, cache=TRUE}
ind <- intersect(rownames(METADATA), colnames(COUNT))
METADATA <- METADATA[ind,]
COUNT <- COUNT[, ind]
PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT,MIN_GENE_CPM=1, MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
```
### Covariate clustering
Determine relationship between covariates. 
```{r covariates.clustering, cache=TRUE}
FactorCovariates <- c('TISSUE','Donor','RACE','CDR','CERJ','SEX','NP.1','bbscore','Dx')
ContCovariates <- c('PMI','AOD','PlaqueMean')

@@

# Add variable of interest back to the residuals
RESIDUAL.GENE_EXPRESSION =

@@

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE}
# Code
CODE <- File(paste0('./',thisFileName), name = "Differential Expression Analysis MSSM RNASeq", parentId = parentId)
CODE <- synStore(CODE, used = ALL_USED_IDs, activityName = activityName, executed = thisFile, activityDescription = activityDescription)

# Fold change
FC <- FC[, unique(colnames(FC))]
write.table(FC, file='./FoldChange_MSSM_RNASeq.tsv', sep='\t', row.names=F, quote=F)
FC_OBJ <- File('./FoldChange_MSSM_RNASeq.tsv',name = 'Differential Expression Analysis MSSM RNASeq logFC',parentId = parentId)
FC_OBJ <- synStore(FC_OBJ, used = ALL_USED_IDs, activityName = activityName, executed=CODE$properties$id, activityDescription = activityDescription)

# Pvalue
PVAL <- PVAL[, unique(colnames(PVAL))]
write.table(PVAL, file='./AdjustedPval_MSSM_RNASeq.tsv', sep='\t', row.names=F, quote=F)
PVAL_OBJ <- File('./AdjustedPval_MSSM_RNASeq.tsv',name = 'Differential Expression Analysis MSSM RNASeq adjPval',parentId = parentId)
PVAL_OBJ <- synStore(PVAL_OBJ, used = ALL_USED_IDs, activityName = activityName, executed=CODE$properties$id, activityDescription = activityDescription)
```
|  *Results*           |  *SynapseID*                    |
  |  -------             |   ---------                     |
  |  Log fold change     |  `r FC_OBJ$properties$id`       |
  |  Adjusted pvalues    |  `r PVAL_OBJ$properties$id`     |

>>>>

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
Data were obtained from AMP-AD portal of MSSM, and following brain regions were analysed for differential expression between case and control:
  | Region | Name                      | 
  | :----- | ------------------------- | 
  | BM_10  | Frontal Pole (FP)         |
  | BM_22  | Superior Temporal Gyrus   |
  | BM_36  | Parahipocampal Gyrus (PHG)|

@@

Remove genes that have less than 1 cpm counts in at least 50% of samples.
```{r cpmnormalisation, cache=TRUE}
ind <- intersect(rownames(METADATA), colnames(COUNT))
METADATA <- METADATA[ind,]
COUNT <- COUNT[1:1000, ind]
PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT,MIN_GENE_CPM=1, MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
```
### Covariate clustering
Determine relationship between covariates. 
```{r covariates.clustering, cache=TRUE}
FactorCovariates <- c('TISSUE','Donor','RACE','CDR','CERJ','SEX','NP.1','bbscore','Dx')
ContCovariates <- c('PMI','AOD','PlaqueMean')

@@

@@

```{r varsToAddBack}
# Add variable of interest back to the residuals
varsToAddIn = c("TISSUE.DxBM_10.AD", "TISSUE.DxBM_10.Control", "TISSUE.DxBM_22.AD",
                "TISSUE.DxBM_22.Control", "TISSUE.DxBM_36.AD","TISSUE.DxBM_36.Control")
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + 
  VOOM.ADJUSTED.FIT$coefficients[,varsToAddIn] %*% t(DM1$design[,varsToAddIn])
```

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE}
# Code
CODE <- Folder(name = "MSBB RNASeq", parentId = parentId)
CODE <- synStore(CODE)

# Expression
expr = rownameToFirstColumn(RESIDUAL.GENE_EXPRESSION, 'EnsembleIDs')
write.table(expr, file='expression.tsv', sep = '\t', row.names=F, quote=F)
obj = File('expression.tsv', 
           name = 'Voom Adjusted Weighted Residual Expression', 
           parentId = parentId)
obj = synStore(obj,
               used = ALL_USED_IDs, 
               activityName = activityName, 
               executed = thisFile, 
               activityDescription = activityDescription)

# Covariates
covar = rownameToFirstColumn(COVARIATES, 'SampleId')
write.table(expr, file='covariates.tsv', sep = '\t', row.names=F, quote=F)
obj = File('covariates.tsv', 
           name = 'Covariates', 
           parentId = parentId)
obj = synStore(obj,
               used = ALL_USED_IDs, 
               activityName = activityName, 
               executed = thisFile, 
               activityDescription = activityDescription)
```
